---
title: "wildlife_corridor.rmd"
author: "Andy Atallah"
date: "2023-11-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
# Install packages
install.packages('tidyverse')
install.packages('here')
install.packages('sf')
install.packages('stars')
install.packages('raster')
library(tidyverse)
library(here)
library(sf)
library(stars)
#library(rgdal)
library(raster)
```


1. Clip landcover by bounding_box.

landcover (GEOMETRY FILE) [CLIP] initial_clip (GEOMETRY FILE)

Inputs: landcover, bounding_box
Outputs: unnamed intermediate "lc_study_clip"
Possible functions: raster::raster

Note: unplanned deviation from original study
Decided not to clip landcover to initial_clip before clipping to bounding_box, as this seemed unnecessary

```{r}
landcover <- here("data", "raw", "public", "landcover.tif") %>% raster()
bounds <- here("data", "raw", "public", "bounding_box.shp") %>% st_read()

st_crs(bounds)
proj4string(landcover)

# Reproject bounds to landcover CRS
bounds <- bounds %>% st_transform(proj4string(landcover))


# Crop
lc_study_clip <- mask(crop(landcover, extent(bounds)), bounds)

```

```{r}
plot(lc_study_clip)
```

2. Reproject landcover to CRS=32736 and 10m resolution.

Unplanned deviation: used a pixel size of 10m (as specified in the manuscript) as opposed to 70m (as specified in the QGIS model).

Inputs: initial_raster_clip
Outputs: unnamed_intermediate "initial_raster_clip_proj"
Possible functions: raster::projectRaster
```{r}

lc_study_clip_reproj <- projectRaster(lc_study_clip, crs=32736)

#Andy's code:
pixel_default <- raster(resolution=c(10,10), crs=proj4string(lc_study_clip_reproj), ext=extent(lc_study_clip_reproj))

lc_study_clip_reproj <- resample(lc_study_clip_reproj, pixel_default)

```

```{r}
plot(lc_study_clip_reproj)
```

3. Reclassify landcover by table.

Unplanned deviation: changed reclass table to intervals

Inputs: "lc_study_clip"
Outputs: unnamed intermediate "lc_reclass"
Possible functions: raster::reclassify
```{r}
# Create 3x10 table of values with which to reclassify
# Data source: the online Google Drive folder flowchart

c1 <- c(0,1,2,3,4,5,6,7,8,9)
c2 <- c(0,2,3,4,5,6,7,8,9,255)
c3 <- c(0,0.75,0.01,0.25,0.01,0.15,1,0.1,0.15,0)
reclass_table <- as.matrix(data.frame(c1, c2, c3))

lc_reclass <- reclassify(lc_study_clip_reproj, reclass_table, right=F)
```

```{r}
plot(lc_reclass)
```


4. Reproject secondary roads to CRS=32736.

secondary_roads (GEOMETRY FILE) --> [Reproject] to EPSG:32736 - WGS 84 / UTM zone 36S

Inputs: secondary_roads
Outputs: unnamed intermediate "output_1a"
Possible functions: sf::read_sf, raster::shapefile, st::st_transform
```{r}
secondary_roads <- here("data", "raw", "public", "secondary_roads.shp") %>% st_read()

output_1a <- st_transform(secondary_roads, crs=32736) %>% st_crop(st_bbox(lc_study_clip_reproj))
```

1b. Rasterize secondary roads.

"1a output" (GEOMETRY FILE) --> [Rasterize]

Inputs: "output_1a"
Outputs: unnamed intermediate "secondary_roads_raster"
Possible functions: stars::st_rasterize
```{r}
secondary_roads_raster <- st_rasterize(output_1a)

#Takes as an input a stars raster layer
#CJ Brown 2020-05-18

st_as_raster <- function(rstars){
  rext <- st_bbox(rstars)
  raster(t(rstars[[1]]), xmn = rext[1], xmx = rext[3],
                   ymn = rext[2], ymx=rext[4],
                   crs = st_crs(rstars)$proj4string)
}

secondary_roads_raster <- st_as_raster(secondary_roads_raster) %>% resample(lc_study_clip_reproj)

secondary_roads_raster <- reclassify(secondary_roads_raster, cbind(NA, 0)) %>% reclassify(cbind(1, Inf, 1))

```

```{r}

plot(secondary_roads_raster)

```

Major Roads

Unplanned deviation: buffer after rasterizing

```{r}
major_roads <- here("data", "raw", "public", "major_roads_vector.shp") %>% st_read()

major_roads_reproj <- major_roads %>% st_transform(crs=32736) %>% st_crop(st_bbox(lc_study_clip_reproj))

major_roads_raster <- st_rasterize(major_roads_reproj)

major_roads_raster <- st_as_raster(major_roads_raster)

major_roads_buff <- major_roads_raster %>% buffer(width=300)

major_roads_raster <- major_roads_buff %>% resample(lc_study_clip_reproj)

major_roads_raster <- reclassify(major_roads_raster, cbind(NA, 0)) %>% reclassify(cbind(1, Inf, 1))

```

```{r}
plot(major_roads_raster)

```

Bomas (rural)

```{r}
rural <- here("data", "raw", "public", "bomas.shp") %>% st_read()

rural_reproj <- rural %>% st_transform(crs=32736)

rural_raster <- st_rasterize(rural_reproj)

rural_raster <- st_as_raster(rural_raster) %>% resample(lc_study_clip_reproj) %>% setExtent(extent(lc_study_clip_reproj))

```

```{r}
plot(rural_raster)

```

Buildings

```{r}
buildings <- here("data", "raw", "public", "merged_buildings.shp") %>% st_read()

buildings_reproj <- buildings %>% st_transform(crs=32736) %>% st_crop(st_bbox(lc_study_clip_reproj))

buildings_raster <- st_rasterize(buildings_reproj)

buildings_raster <- st_as_raster(buildings_raster) %>% resample(lc_study_clip_reproj)

```

```{r}
plot(buildings_raster)

```

Figure out why buildings are not rasterizing properly

```{r}

merged_buildings <- overlay(buildings_raster, rural_raster, fun=function(x,y){return(x | y)}) %>% reclassify(cbind(1, Inf, 1))

```

```{r}
plot(merged_buildings)

```


```{r}
proximity = distance(merged_buildings)

plot(proximity)

```

Reclassify proximity raster

```{r}
c1 <- c(0, 1, 50, 100, 150, 200)
c2 <- c(1, 50, 100, 150, 200, Inf)
c3 <- c(1, 0.8, 0.6, 0.4, 0.2, 0)
prox_reclass_table <- as.matrix(data.frame(c1, c2, c3))

prox_reclass <- reclassify(proximity, prox_reclass_table)
```

```{r}
plot(prox_reclass)

```


6a. Rasterize buffer zone.

Inputs: "buffer_region"
Outputs: unnamed intermediate "buffer_raster"
Possible functions: raster::raster
```{r}
buffer_zone <- here("data", "raw", "public", "buffer_region.shp") %>% st_read()

buffer_reproj <- buffer_zone %>% st_transform(crs=32736) %>% st_crop(st_bbox(lc_study_clip_reproj))

buffer_raster <- st_rasterize(buffer_reproj)

buffer_raster <- st_as_raster(buffer_raster) %>% resample(lc_study_clip_reproj)

buffer_raster <- reclassify(buffer_raster, cbind(1, Inf, 999)) %>% reclassify(cbind(NA, 0))

plot(buffer_raster)

```

(A+(0.25*B)+C+D+E)/4
A = major roads
B = secondary roads
C = proximity to buildings
D = landcover
E = buffer zone

```{r}
A = major_roads_raster
B = secondary_roads_raster
C = prox_reclass
D = lc_reclass
E = buffer_raster

cost <- (A+(0.25*B)+C+D+E)/4

plot(cost)

```

Issues:

- buildings don't rasterize
- secondary roads are too thick (70m resolution)
- should everything be in 10m resolution?
- workflow doesn't show accumulated cost

Compare to cost surface in the google drive (subtract)

```{r}
compare <- here("data", "derived", "public", "DD_Cost_6.tif") %>% raster()

compare <- compare %>% resample(pixel_default)
cost <- cost %>% resample(pixel_default)

diff <- overlay(compare,cost,
                      fun=function(compare, cost){return(compare-cost)}, filename="difference")
plot(diff)
```

Differences:

- buildings (because we couldn't rasterize them)
- secondary roads are too wide
